---
title: "Project"
author: "Vicki M. Zhang"
date: "January 29, 2020"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = F, message = F, warning = F,
                      fig.width = 6, fig.height = 6)

library(rethinking)
library(dplyr)
```

```{r}
t_dam <- read.csv("taraxacum_damage_2018.csv")
```
## Visualizing all data

```{r}
str(t_dam)
```

Description:

- 180 rows of data
- columns:
    1. "plant" = plant ID (1 to 180)
    2. "date" = date of data collection
    3. "site" = site of data collection (1-5)
    4. "town" = binary Y/N of whether plants were found in Churchill townsite
    5. "species" = three *Taraxacum* species (*T. officinale* is invasive, *T. lacerum* and *T. ceratophorum* are native)
    6. "leaf1", "leaf2", "leaf3" = measures of herbivory of 3 oldest leaves (i.e., largest, closest to ground)
  

Histogram
```{r}
t_dam %>% 
  ggplot(aes(x = site, fill = species)) + 
  geom_histogram() +
  theme_classic() +
  labs(title = "Distribution of Individuals across Sites",
       x = "Sites",
       y = "Number of Individuals")
```

## Species-Level Differences of Herbivory

__Number of individuals per site__
```{r}
t_dam %>% 
  group_by(site, town, species) %>% 
  tally()
```

Notes:

* sites 1-3 were in town, sites 4-5 were not in town
* data for all three species were only collected in site 2
    + only data for *T. ceratophorum* in site 2
* sites 1 and 3 have data for only *T. lacerum* and *T. officinale*
* sites 4 and 5 (i.e., out of town sites) have data only for *T. lacerum*

__Number of individuals per species__
```{r}
t_dam %>% 
  group_by(species) %>% 
  tally()
```

N =

* *T. ceratophorum* = 20
* *T. lacerum* = 100
* *T. officinale* = 60

__Averaged damage of all plants in three species__
```{r}
t_dam %>% 
  group_by(species) %>% 
  summarize(leaf1 = mean(leaf1),
            leaf2 = mean(leaf2),
            leaf3 = mean(leaf3),
            average = ((leaf1 + leaf2 + leaf3) / 3))
```

Table:

* "leaf1", "leaf2" and "leaf3" columns are averaged damage percentages for each leaf
* "average" is the average of columns "leaf1", "leaf2" and "leaf3"



```{r}
# Distribution of damage across species
t_dam %>% 
  mutate(avg_dam = ((leaf1 + leaf2 + leaf3) / 3)) %>% 
  ggplot(aes(x = species, y = avg_dam)) +
  geom_boxplot() +
  labs(title = "Average damage across species",
       x = "Species",
       y = "Damage (averaged across 3 largest leaves)") +
  theme_classic()
```

Notes:

* *T. officinale* (the invasive species) has the greatest herbivory damage
* no general pattern over which leaf is most damaged by herbivory
    + dandelion leaves are in a floret near the base of the plant
    + since all the leaves are close to each other, the lack of difference makes biological sense


__Number of plants per species with no damage__
```{r}
t_dam %>%
  group_by(species) %>% 
  filter(leaf1 == "0" & leaf2 == "0" & leaf3 == "0") %>% 
  tally()
```

Notes:

* *T. ceratophorum* = 20
    + 3 individuals with no damage
* *T. lacerum* = 100
    + 61 individuals with no damage
* *T. officinale* = 60
    + 24 individuals with no damage
  
__Averaged damage of all plants minus plants with no damage__
```{r}
t_dam %>% 
  group_by(species) %>% 
  filter(leaf1 != "0" & leaf2 != "0" & leaf3 != "0") %>%
  summarize(leaf1 = mean(leaf1),
            leaf2 = mean(leaf2),
            leaf3 = mean(leaf3),
            average = ((leaf1 + leaf2 + leaf3) / 3))  
```


Table:

* "leaf1", "leaf2" and "leaf3" columns are averaged damage percentages for each leaf
* "average" is the average of columns "leaf1", "leaf2" and "leaf3"
* this only includes plants with damage (filtered out plants with no damage)



```{r}
# Distribution of damage across species minus plants with no damage
t_dam %>%
  filter(leaf1 != "0" & leaf2 != "0" & leaf3 != "0") %>%
  mutate(avg_dam = ((leaf1 + leaf2 + leaf3) / 3)) %>% 
  ggplot(aes(x = species, y = avg_dam)) +
  geom_boxplot() +
  labs(title = "Average damage across species (minus plants with no damage)",
       x = "Species",
       y = "Damage (averaged across 3 largest leaves)") +
  theme_classic()
```

Notes:

* *T. officinale* (the invasive species) again has the greatest herbivory damage
* no general pattern over which leaf is most damaged by herbivory

## Differences in Herbivory between sites

__Average damage in and out of town__
```{r}
t_dam %>% 
  group_by(town, species) %>% 
  summarize(leaf1 = mean(leaf1),
            leaf2 = mean(leaf2),
            leaf3 = mean(leaf3),
            average = ((leaf1 + leaf2 + leaf3) / 3)) 
```

```{r}
# Distribution of average damage across species
t_dam %>%
  group_by(town, species) %>% 
  mutate(avg_dam = ((leaf1 + leaf2 + leaf3) / 3)) %>% 
  ggplot(aes(x = species, y = avg_dam)) + facet_wrap(~ town) +
  geom_boxplot() +
  labs(title = "Average damage in and out of town",
       x = "Species",
       y = "Damage (averaged across 3 largest leaves)") +
  theme_classic()
```

Notes:

* *T. lacerum* was the only species that had data both in and out of town
* more damage in town



__Average damage across species in town__ 
```{r}
t_dam %>% 
  group_by(town, species) %>% 
  filter(town == "Y") %>% 
  summarize(leaf1 = mean(leaf1),
            leaf2 = mean(leaf2),
            leaf3 = mean(leaf3),
            average = ((leaf1 + leaf2 + leaf3) / 3)) 
```


```{r}
# Distribution of average damage across species in town
t_dam %>%
  group_by(town, species) %>% 
  filter(town == "Y") %>% 
  mutate(avg_dam = ((leaf1 + leaf2 + leaf3) / 3)) %>% 
  ggplot(aes(x = species, y = avg_dam)) + facet_wrap(~ site) +
  geom_boxplot() +
  labs(title = "Average damage across species in town",
       x = "Species",
       y = "Damage (averaged across 3 largest leaves)") +
  theme_classic() + theme(axis.text.x = element_text(angle = 90))
```


Notes:

* *T. officinale* (invader) had  most herbivory damage
* more damage in town

## Effects of Date of Data Collection

__Date of data collection across sites__
```{r}
t_dam %>% 
   group_by(date, site) %>% 
   tally()
```

Notes

- most of the data (N = 133) collected on August 27th
      * data collected from sites 1-4
- some data collected Aug 28 (N = 46) and Aug 30 (N = 1)
      * data collected on Aug 28th from sites 2, 3, 5
      * one data point collected Aug 30th from site 3


```{r}
t_dam %>%
   ggplot(aes(x = site, fill = date)) +
   geom_bar() +
   theme_classic() +
   labs(title = "Number of data points collected per site across date",
        x = "Site",
        y = "N")
```

__Date of data collection across species__

```{r}
t_dam %>% 
   group_by(date, species) %>% 
   tally()
```

Notes

- Aug 27th data (N = 133)
      - only *T. lacerum* and *T. officinale* data collected
- Aug 28th data (N = 46)
      - data from all three species collected
      - N (*T. ceratophorum*) = 20
      - N (*T. officinale*, the invader) = 2
      - N (*T. lacerum*) = 24
- Aug 29th only has one data point for *T. lacerum*

```{r}
t_dam %>%
   ggplot(aes(x = site, fill = species)) + facet_grid(~ date) +
   geom_bar() +
   theme_classic() +
   labs(title = "Number of species data collected per site across date",
        x = "Site",
        y = "N")
```

## General thoughts

1. The variables that are affecting the system here are:

   - site (1-5)
   - in or out of town (Y/N)
   - species


2. I want to predict herbivory damage

   - not sure if I should use all three herbivory measures, choose one, or average the three measurements into a new measure
   - also, not sure if species is a predictor, or if I should predict herbivory damage for all three species




unsure variables

- site affects herbivory damage
- not really sure how site and town are connected




















