---
title: "Project"
author: "Vicki M. Zhang"
date: "March 30, 2020"
output: 
   html_document:
      toc: true
      toc_depth: 4
      toc_float: true
      df_print: paged
      code_folding: hide
---

# Introduction

## Setting Up

```{r setup, include = F}
knitr::opts_chunk$set(message = F, warning = F)

library(rethinking)
library(tidyverse)
library(dagitty)
```


## Data Description

```{r}
d <- read.csv("taraxacum_damage_2018.csv")
str(d)
```

This dataset contains herbivore damage information of three species of Taraxacum: *T. officinale*, a widely-invasive dandelion, *T. lacerum*, the native and common dandelion, and *T. ceratophorum*, a native but more rare dandelion. Data was collected from five sites, three of which were within the Churchill townsite footprint, and the other two in the surrounding natural areas. However, herbivory data was not collected equally between species and between sites. *T. lacerum* data (N = 100) was collected from all five sites, while *T. officinale* (N = 60) was only found within Churchill, and thus data was only collected from sites 1, 2, and 3. *T. ceratophorum* data (N = 20) was only collected in site 2 within town.

Sites:

- Site 1: Manitoba Port sign (N edge of town)
- Site 2: Town Square
- Site 3: Town Complex (Hospital and school)
- Site 4: Southern RR line on Goose Creek Road
- Site 5: Goose Creek Pumping Station

In-town sites (1, 2, 3) are arranged in order of least disturbed (1) to most disturbed (3).

```{r}
d %>% 
   group_by(site, town, species) %>% 
   tally()
```


The measures of herbivore damage are taken from the three largest leaves, which are also the three oldest leaves and are closest to the soil. However, since dandelions grow from a rosette, these three leaves may not differ greatly in terms of herbivore damage compared to other leaves around it. The table below displays the average percentage of herbivore damage for the three leaves in each species, as well as an averaged total damage across all three leaves.

```{r}
d <- d %>% 
   dplyr::select(-plant, -date) %>% 
   group_by(site, species) %>% 
   mutate(avg = (leaf1 + leaf2 + leaf3) / 3)


d %>% 
   summarize(avf_leaf1 = mean(leaf1),
             avg_leaf2 = mean(leaf2),
             avg_leaf3 = mean(leaf3),
             avg_total = mean(avg))
```

My model will aim to predict herbivore damage for *T. officinale* and *T. lacerum* in in-town sites (sites 1-3)

First, I ensured that all measures of herbivory are numeric, and `site` are integers
```{r}
# as.numeric
d$leaf1 <- as.numeric(d$leaf1)
d$leaf2 <- as.numeric(d$leaf2)
d$leaf3 <- as.numeric(d$leaf3)

# as.numeric for site
d$site <- as.integer(d$site)

# check average herbivory measures
d %>% 
   group_by(site, species) %>% 
   summarize(avg_leaf1 = mean(leaf1),
             avg_leaf2 = mean(leaf2),
             avg_leaf3 = mean(leaf3),
             avg_total = mean(avg))

```

Create a new data frame `d_town` with only the observations in town.

```{r}
# data frame with only in-town observations
d_town <- d %>%
   mutate(avg = (leaf1 + leaf2 + leaf3) / 3) %>% 
   dplyr::select(site, species, leaf1, leaf2, leaf3, avg) %>% 
   filter(site == 1 |
             site == 2 |
             site == 3)

```



Create a new data frame `d_lac` with only _T. lacerum_ data. It might be interesting to just look at this one common native species, eventually.

```{r}
# data frame with only T. lacerum
d_lac <- d %>% 
   filter(species == "T_lacerum") %>% 
   dplyr::select(site, town, leaf, herb)
```



In `d_town`, species is also a categorical variable. Here, I am indexing the species so that *T. lacerum* is indexed as `1` and *T. officinale* is indexed as `2`.
```{r}
d_town$species <- ifelse(d_town$species == "T_lacerum", 1, 2)
str(d_town)
```

Summary:
```{r}
precis(d)
precis(d_cerat)
precis(d_town)
```

# Model 1: Predicting Herbivory

Plot `leaf1`, `leaf2`, `leaf3` and `avg` (average herbivory) by `species`, where `1` is *T. lacerum* and `2` is *T. officinale*.
```{r}
plot(d_town$leaf1 ~ d_town$species)
plot(d_town$leaf2 ~ d_town$species)
plot(d_town$leaf3 ~ d_town$species)
plot(d_town$avg ~ d_town$species)
```

- there may be a relationship here, where *T. lacerum* has less herbivory damage than *T. officinale*
- one *T. lacerum* outlier


## Variables and DAG

The model will use the variables of `species`, `leaf1`, `leaf2` and `leaf3` to predict total herbivory in town. Different sites `site` and other unknown variables (U) may affect herbivore damage as well.

Below is the DAG for this model. I am predicting total herbivory using the average herbivore damage measurement (avg), and species (sp). Unknown variables (U), such as site, may affect herbivore damage as well. 

```{r}
d_dag <- dagitty("dag {
            sp -> HD 
            avg -> HD 
            sp -> avg 
            U -> HD
            } ")

coordinates(d_dag) <-  list(x = c(sp = 0, HD = 1, avg = 2, U = 1),
                                y = c(sp = 0, HD = 1, avg = 0, U = 2))
drawdag(d_dag)
```

Variables

   - `species` is categorical: *T. lacerum*, *T. officinale*
   - `leaf1`, `leaf2` and `leaf3` are continuous (percentages) but are low
   - `site` is an integer: 1, 2, 3


Distribution of `avg`
```{r}
dens(d_town$avg)
```

- not Gaussian distribution but let's just keep going for now
- empirical distribution doesn't need to be Gaussin in order to justify using a Guassian probability distribution



## Linear Model using normal prior

Linear model of average herbivory `avg` by `species`:

- `1` indexes to *T. lacerum* and `2` indexes to *T. officinale* (no order)
- two $\alpha$ parameters, where `j` corresponds to species

First pass at using binomial priors, using both `quap` and `ulam`

$$avg_i \sim Binomial(\mu_i, \sigma)$$
$$\mu_i = \alpha_{species[i]} + \beta_s site[i]$$

$$\alpha_i \sim Normal(1, 0.1)$$

$$\beta_i \sim Normal(0, 10)$$

$$\sigma \sim Uniform(2, 5)$$


### Priors

Normal(1, 0.25) prior for $\alpha$
```{r}
curve(dnorm(x, 1, 0.1), from = 0, to = 100)
```

- mean of the distribution is around 1 (i.e., average of 1% herbivory)

Normal(1, 3) prior for $\beta$
```{r}
curve(dnorm(x, 1, 3), from = -10, to = 50)
```



```{r}
a <- dnorm(1e4, 1, 0.1)
b <- rnorm(1e4, 1, 3)

plot(NULL, xlim = range(d_town$species), ylim = c(-10, 50),
     xlab = "species", ylab = "average herbivory %")
abline(h = 0, lty = 2)

x <- d_town$site

mtext("b ~ dnorm(1, 3)")
for(i in 1:N)
  curve(a[1] + b[1] * x,
        from = 1, to = 2, add = T,
        col = col.alpha("black", 0.2))

```

Normal(0, 3) prior for $\beta$
```{r}
curve(dnorm(x, 0, 3), from = 0, to = 100)

```


```{r}
a <- rnorm(1e4, 1, 0.1)
b <- rnorm(1e4, 0, 3)

plot(NULL, xlim = range(d_town$species), ylim = c(-10, 50),
     xlab = "species", ylab = "average herbivory %")
abline(h = 0, lty = 2)

x <- d_town$species

mtext("b ~ dnorm(0, 10)")
for(i in 1:N)
  curve(a[1] + b[1] * x,
        from = 1, to = 2, add = T,
        col = col.alpha("black", 0.2))

```



Uniform prior for $\sigma$ 
```{r}
curve(dunif(x, 2, 10), from = 0, to = 15)
```




```{r}
N <- 100
a_bar <- rlnorm(1, 0.5)
sigma_a <- exp(1)
a <- rlnorm(N, a_bar, sigma_a)
b_bar <- rlnorm(0, 1.5)
b <- rlnorm(N, b_bar, 1.5)
sigma_g <- exp(1)
g <- rlnorm(N, 0, sigma_g)


plot(NULL, xlim = range(d_town_slim$species), ylim = c(-5, 100),
     xlab = "species", ylab = "herb")
for(i in 1:N)
   curve(a[i] + b[i]*x + c[i],
         from = 1, to = 2, add = T,
         col = col.alpha("black", 0.2))
```

### quap1 model

```{r}
m1q <- quap(
   alist(
      avg ~ dbinom(mu, sigma),
      mu <- a[species] + b[site],
      a[species] ~ dnorm(1, 0.1),
      b[site] ~ dnorm(0, 3),
      sigma ~ dunif(2, 10)
   ), data = d_town
)

precis(m1q, depth = 2)
```

- species `2`, *T. officinale*, has a higher herbivore damage percentage
- the more disturbed sites have greater herbivory
- did not capture the interaction between `species` and `site`



#### Posterior inference

I can't figure out how to plot the posterior inference against the data. Since both of my predictor variables are not continuous, I'm not really sure how the implied regression line should look.


```{r}
prior_m1q <- extract.prior(m1q)
```

```{r}
plot(NULL, xlim = c(1, 2), ylim = c(-5, 15),
     xlab = "species", ylab = "herbivory damage %")
abline(h = 0)

m1q_seq <- seq(from = 0, to = 2, length.out = 30)
mu <- link(m1q, post = prior_m1q, data = data.frame(species = m1q_seq))
for(i in 1:50)
   lines(m1q_seq, mu[i, ], col = col.alpha("black", 0.3))
```


I also tried to use only a sample of the data to see the scatter. First, I sampled N number of rows from the data frame:

```{r}
set.seed(12)
N <- 50
dN <- d_town[sample(nrow(d_town), N),]
head(dN)
```

Re-estimating the model:

```{r}
mN <- quap(
   alist(
      avg ~ dnorm(mu, sigma),
      mu <- a[species] + b[site],
      a[species] ~ dnorm(1, 0.1),
      b[site] ~ dnorm(0, 3),
      sigma ~ dunif(2, 10)
   ), data = dN
)

precis(mN, depth = 2)
```


Using the `link` function:

```{r}
mu <- link(m1q)
str(mu)
```

### quap2 model


```{r}
m2q <- quap(
   alist(
      avg ~ dnorm(mu, sigma),
      mu <- a[species] + b[site],
      a[species] ~ dnorm(1, 0.1),
      b[site] ~ dnorm(0, 3),
      sigma ~ dunif(2, 10)
   ), data = d_town
)

precis(m2q, depth = 2)
```

#### Posterior inference

```{r}
plot(avg ~ species, data = d_town, col = rangi2)

post_m1q <- extract.samples(m1q)
a_map <- mean(post_m1q$a[species])
b_map <- mean(post_m1q$b[site])
curve(a_map + (b_map * x), add = TRUE)
```

```{r}
N <- 50
dN <- d_town[sample(nrow(d_town), N),]

mN <- quap(
   alist(
      avg ~ dnorm(mu, sigma),
      mu <- a[species] + bS * site,
      a[species] ~ dnorm(1, 0.1),
      bS ~ dnorm(0, 3),
      sigma ~ dunif(2, 10)
   ), data = dN
)

precis(mN, depth = 2)

# plot 20 of these lines using model
post <- extract.samples(mN, n = 20)

plot(dN$species, dN$avg,
     xlim = range(d_town$species), ylim = range(d_town$avg),
     col = rangi2, xlab = "species", ylab = "average herbivory %")
mtext(concat("N = ", N))

for(i in 1:20)
   curve(post$a[1] + post$bS[1] * x,
         col = col.alpha("black", 0.3), add = T)
```

below doesn't work

```{r}
plot(NULL, xlim = c(1, 2), ylim = c(-5, 10),
     xlab = "species", ylab = "herbivory damage %")
abline(h = 0)

m1q_seq <- seq(from = 0, to = 2, length.out = 30)
mu <- link(m1q, post = prior_m1q, data = data.frame(species = m1q_seq))
for(i in 1:50)
   lines(m1q_seq, mu[i, ], col = col.alpha("black", 0.3))
```



- it appears that *T. lacerum* has less herbivory damage than *T. officinale*

```{r}
plot(avg ~ species, data = d_town, col = rangi2)
map <- mean(post_m1q$avg)
curve(map, add = TRUE)

post_m1q <- extract.samples(m1q)
```

### quap3 model

```{r}
m3q <- quap(
   alist(
      avg ~ dnorm(mu, sigma),
      mu <- a[species] + bS*site,
      a[species] ~ dnorm(1, 0.1),
      bS ~ dnorm(0, 3),
      sigma ~ dunif(2, 10)
   ), data = d_town
)

precis(m1q, depth = 2)
```


### which models?

#### Prior predictive simulation

1st simulation of plant herbivory by sampling from posterior
```{r}
sample_mu <- rnorm(1e4, 1, 0.1)
sample_sigma <- runif(1e4, 2, 5)
prior_avg <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior_avg)
```

- can't have negative herbivory, which this model says is possible



1st simulation of plant herbivory by sampling from posterior
```{r}
sample_mu <- rnorm(1e4, 1, 0.1)
sample_sigma <- runif(1e4, 0, 0.25)
prior_avg <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior_avg)
```



First pass at priors for `a` and `b`. Simulating lines:
```{r}
# simulating lines
set.seed(1)
N <- 100
a <- rnorm(N, 1, 0.1)
b <- rnorm(N, 0, 0.25)
dens(b, xlim = c(0, 1), adj = 0.1)


# plotting simulated lines
plot(NULL, xlim = range(d_town$species), ylim = c(-5, 10),
     xlab = "species", ylab = "average herbivory damage (%)")
abline(h = 0, lty = 2)
mtext("b ~ dnorm(0, 0.25)")

xbar <- mean(d_town$avg)
for(i in 1:N)
   curve(a[i] + b[i] * (x - xbar),
         from = 1, to = 2, add = TRUE,
         col = col.alpha("black", 0.2))
```

- this doesn't look too bad?



Second pass: defining `b` as Log-Normal:
$$\beta \sim Log-Normal(0, 1)$$


```{r}
set.seed(1)
N <- 100
a <- rnorm(N, 1, 0.1)
b <- rlnorm(N, 0, 0.5)
dens(b, xlim = c(0, 5), adj = 0.1)

# plotting 2nd simulated lines

plot(NULL, xlim = range(d_town$species), ylim = c(-5, 10),
     xlab = "species", ylab = "average herbivory damage (%)")
abline(h = 0, lty = 2)
mtext("b ~ rlnorm(0, 0.25)")

xbar <- mean(d_town$avg)
for(i in 1:N)
   curve(a[i] + b[i] * (x - xbar),
         from = 1, to = 2, add = TRUE,
         col = col.alpha("black", 0.2))
```

- not really sure which one is better here, normal or log-normal





### ulam with logit



```{r}
m3q <- ulam(
   alist(
      species ~ dbinom(1, p),
      logit(p) <- avg,
      avg ~ dnorm(1, 0.1)
   ), data = d_town
)
```





## Linear Model using log-normal prior

$$avg_i \sim Normal(\mu_i, \sigma)$$
$$\mu_i = \alpha_{species[i]} + \beta_s site[i]$$

$$\alpha_i \sim Normal(1, 0.1)$$

$$\beta_i \sim Normal(0, 10)$$

$$\sigma \sim Log-Normal(2, 5)$$

Second pass: defining `b` as Log-Normal:
$$\beta \sim Log-Normal(0, 1)$$


```{r}
set.seed(1)
N <- 100
a <- rnorm(N, 1, 0.1)
b <- rlnorm(N, 0, 0.5)
dens(b, xlim = c(0, 5), adj = 0.1)

# plotting 2nd simulated lines

plot(NULL, xlim = range(d_town$species), ylim = c(-5, 10),
     xlab = "species", ylab = "average herbivory damage (%)")
abline(h = 0, lty = 2)
mtext("b ~ rlnorm(0, 0.25)")

xbar <- mean(d_town$avg)
for(i in 1:N)
   curve(a[i] + b[i] * (x - xbar),
         from = 1, to = 2, add = TRUE,
         col = col.alpha("black", 0.2))
```

- not really sure which one is better here, normal or log-normal




### Priors

Plotting the $\sigma$ prior
```{r}
curve(dlnorm(x, 2, 10), from = 0, to = 15)
```


Linear model of average herbivory `avg` by `species`:

- `1` indexes to *T. lacerum* and `2` indexes to *T. officinale* (no order)
- two $\alpha$ parameters, where `i` corresponds to species


## quap model

```{r}
m2q <- quap(
   alist(
      avg ~ dnorm(mu, sigma),
      mu <- a[species] + b[site],
      a[species] ~ dnorm(1, 0.1),
      b[site] ~ dnorm(0, 3),
      sigma ~ dlnorm(2, 10)
   ), data = d_town
)

precis(m2q, depth = 2)
```

### Posterior inference

Again, can't really figure out how to sample from the posterior.


## ulam

```{r}
m2 <- ulam(
   alist(
      avg ~ dnorm(mu, sigma),
      mu <- a[species] + b[site],
      a[species] ~ dnorm(1, 0.1),
      b[site] ~ dnorm(0, 3),
      sigma ~ dlnorm(2, 10)
   ), data = d_town_slim, chains = 1, cores = 4
)

show(m2)
```

- not really sure how many chains I should use

```{r}
precis(m2, depth = 2)
```

```{r}
pairs(m2)
```

```{r}
traceplot(m2)
trankplot(m2, n_cols = 2)
```


### PPC `m1`


Extract posterior.

```{r}
post_m1 <- extract.samples(m1.2)
p_m1 <- inv_logit(post_m1$b)
precis_plot(precis(as.data.frame(post_m1)))
```


```{r}
m1_fit <- m1.2@stanfit
posterior_m1 <- as.matrix(m1_fit)
dim(posterior_m1)
```



```{r}
tail(colnames(posterior_m1))
hist(posterior_m1[,"b[1]"])
```

```{r}
params <- c(names(m1_fit)[grep("a", names(m1_fit))],
            names(m1_fit)[grep("b", names(m1_fit))])

mcmc_areas(
  posterior_m1, 
  pars = params,
  prob = 0.8,
  prob_outer = 0.999,
  point_est = "mean"
)
```

### Posterior predictive checks

```{r}
y <- d_town_slim$herb

# generate posterior predictions with "sim" in rethinking package
y_rep1 <- sim(m1.2, n = 500)
```

This is just one of the plots you can make, but it's a ggplot object, so you can easily adjust it like I have done here with the rescaling.

```{r}
ppc_dens_overlay(y, y_rep1[1:50, ]) + scale_x_continuous(trans="log1p")
ppc_freqpoly_grouped(y, y_rep1[1:2, ],
                     group = d_town_slim$species,
                     freq = FALSE) + 
   yaxis_text(on = TRUE)
```







## `ulam` with without hierarchical effect of `site`

The model, in mathematical form:

$$avg \sim Poisson(leaf1, p_i)$$
$$logit(p_i) = \alpha_{ind[i]} + \beta_{species[i]}$$
$$\alpha_j \sim Normal(\overline \alpha, \sigma) for j = 1, 2, ... 120$$
$$\overline \alpha \sim Normal(1, 0.25)$$

$$\sigma_{\alpha} \sim Uniform(2, 10)$$
$$\beta_j \sim Normal(0, 3) for j = 1, 2$$



Approximate posterior
```{r}
set.seed(9)
m4 <- ulam(
   alist(
      # regression
      herb ~ dpois(lambda),
      logit(lambda) <- a[ind] + b[species],
      
      # beta prior
      b[species] ~ dnorm(0, 3),
      
      # adaptive priors
      a[ind] ~ dnorm(a_bar, sigma_a),
      
      # hyper-priors
      a_bar ~ dnorm(1, 0.25),
      sigma_a ~ dunif(2, 10)
   ), data = d_town_slim, log_lik = T
)
```


```{r}
precis(m4, depth = 2)
plot(precis(m4, depth = 2))
```

```{r}
compare(m2, m4)
```

## `ulam` with hierarchical effect of `site` and `species`

The model, in mathematical form:

$$avg \sim Poisson(leaf1, p_i)$$
$$logit(p_i) = \alpha_{ind[i]} + \beta_{species[i]} + \gamma_{{site}[i]}$$
$$\alpha_j \sim Normal(\overline \alpha, \sigma) for j = 1, 2, ... 120$$
$$\overline \alpha \sim Normal(1, 0.25)$$

$$\sigma_{\alpha} \sim Uniform(2, 10)$$
$$\beta_j \sim Normal(0, 3) for j = 1, 2$$

$$\gamma_j \sim Normal(0, \sigma_{\gamma}) for j = 1, 2, 3$$
$$\sigma_{\gamma} \sim exp(1)$$



Approximate posterior
```{r}
set.seed(9)
m5 <- ulam(
   alist(
      # regression
      herb ~ dpois(lambda),
      logit(lambda) <- a[ind] + b[species] + g[site],
      
      # adaptive priors
      a[ind] ~ dnorm(a_bar, sigma_a),
      b[species] ~ dnorm(0, sigma_b),
      g[site] ~ dnorm(0, sigma_g),
      
      # hyper-priors
      a_bar ~ dnorm(1, 0.25),
      sigma_a ~ dunif(2, 10),
      sigma_b ~ dexp(1),
      sigma_g ~ dexp(1)
   ), data = d_town_slim, log_lik = T
)
```


```{r}
precis(m5, depth = 2)
plot(precis(m5, depth = 2))
```


```{r}
coeftab(m2, m5)
```



## another `ulam`?

A second model with the interaction between `species` (species ID) and `site`. There's no $\sigma$ prior here, should I include one? idk?

```{r}
m4 <- ulam(
   alist(
      avg ~ dpois(lambda),
      logit(lambda) <- a[species] + b[species] * site,
      a[species] ~ dnorm(1, 0.1),
      b[species] ~ dnorm(0, 3)
   ), data = d_town_slim, chains = 4, log_lik = TRUE
)

show(m4)
precis(m4, depth = 2)
```

- not really sure how many chains I should use

```{r}
precis(m4, depth = 2)
```

```{r}
pairs(m4)
```

```{r}
traceplot(m4)
trankplot(m4, n_cols = 2)
```


```{r}
plot_m7 <- mcmc_combo(m7_fit,
           combo = c("dens_overlay", "trace"),
           pars = c("sigma_g", "sigma_d"),
           transformations = "log1p",
           gg_theme = ggplot2::theme_classic() + legend_none())


plot_m7.1 <- mcmc_combo(m7.1_fit,
           combo = c("dens_overlay", "trace"),
           pars = c("sigma_g"),
           transformations = "log1p",
           gg_theme = ggplot2::theme_classic() + legend_none())

plot_m7.2 <- mcmc_combo(m7.2_fit,
           combo = c("dens_overlay", "trace"),
           pars = c("sigma_d"),
           transformations = "log1p",
           gg_theme = ggplot2::theme_classic() + legend_none())
```



# Model 2: Predicting in-town and out-of-town herbivory of *T. lacerum* 

## Variables and DAG


- whether individuals are in/out of town will be affected herbivore damage
- species will have different herbivory damage
- in/out of town affects species
- also unknown variables!!!



# MESSY CODE BELOW

# Posterior predictions `m1`

## Plotting regression intervals and contours


Creating a sequence of 100 `species`:

```{r}
n <- 1e4
sp_seq <- rep(1:2, length.out = n)
```

There are 100 columns in `lambda` for the 100 simulated values

Using `link`:

```{r}
lambda <- link(m1, data = data.frame(species = sp_seq))
str(lambda)
```

Plot distribution of $\lambda$ at each `species`

```{r}
plot(herb ~ species, d_town)

for(i in 1:100)
   points(sp_seq, lambda[i, ], pch = 16, col = col.alpha(rangi2, 0.1))
```


Use `apply` to summarize distribution of each `species`

```{r}
lmu <- apply(lambda, 2, mean)
lci <- apply(lambda, 2, PI, prob = 0.89)
```



## Prediction Intervals

```{r}
post_m1 <- extract.samples(m1)
p_m1 <- inv_logit(post_m1$a)
precis_plot(precis(as.data.frame(p_m1)))
```


```{r}
sim.sp <- sim(m1, data = list(species = sp_seq))
str(sim.sp)
```

```{r}
herb.PI <- apply(sim.sp, 2, PI, prob = 0.89)
```

```{r}
plot(herb ~ species, d_town_slim, col = col.alpha(rangi2, 0.5))

lines(sp_seq)

shade(herb.PI, sp_seq)
```


```{r}
pairs(m1)
```




# Posterior predictions

## `m2` and `m2.2`
```{r}
precis(post_m2, depth = 2)
str(post_m2)

a_map <- mean(post_m2$a)
b_map <- mean(post_m2$b)

a_map
b_map
```

```{r}
post_m2$diff_sp <- post_m2$b[ ,1] - post_m2$b[, 2]
precis(post_m2, depth = 2)
```

The calculation `diff_sp` is the expected difference in herbivory (out of 100) between *T. lacerum* and *T. officinale*.

```{r}
# postcheck(m2)
```



```{r}
link(m2)
```


```{r}
k <- PSIS(m2, pointwise = TRUE)$K
plot(d_town$species, d_town$herb,
     xlab = "species", ylab = "herbivory %",
     col = rangi2, pch = ifelse(d_town$species == 1, 1, 16), lwd = 2,
     ylim = c(0, 50), cex = 1 + normalize(k))

ns <- 100
site_seq <- seq(from = 1, to = 3, length.out = ns)

# predictions for species 1
lambda <- link(m2, data = data.frame(species = sp_seq, site = 1))
lmu <- apply(lambda, 2, mean)
lci <- apply(lambda, 2, PI)
lines(site_seq, lmu, lty = 2, lwd = 1.5)
shade(lci, site_seq, xpd = TRUE)

# predictions for species 2
lambda <- link(m2, data = data.frame(species = sp_seq, site = 2))
lmu <- apply(lambda, 2, mean)
lci <- apply(lambda, 2, PI)
lines(site_seq, lmu, lty = 3, lwd = 1.5)
shade(lci, site_seq, xpd = TRUE)
```


```{r}
pairs(m2)
```

## `m2.4`

The `link` function takes `ulam` approximation, sample from posterior distribution, and compute $\mu$ for each case in data, and sample from the posterior.

```{r}
mu <- link(m2.4)
str(mu)
```

Each row is a sample from the posterior distribution, each column is a case (row) in the data.

Now, I am calculating the distribution of $\mu$ for each unique species and leaf of each individual.

```{r}
# define sequence of `species` and `leaf` to compute predictions for
n <- 100
sp_seq <- rep(1:2, length.out = ns)
leaf_seq <- rep(1:3, length.out = ns)
```


Plotting this all together

```{r}
plot(d_town_slim$species, d_town_slim$herb,
     xlab = "species", ylab = "herbivory",
     col = rangi2, pch = ifelse(d_town_slim$site == 1, 1,
                                ifelse(d_town_slim$site == 2, 2, 4)),
     ylim = c(0, 60))


# predictions for site = 1
mu <- link(m2.4, data = data.frame(species = sp_seq, leaf = leaf_seq))
mu_mean <- apply(mu, 2, mean)
mu_pi <- apply(mu, 2, PI, prob = 0.89)
lines(sp_seq, mu_mean, lty = 2, lwd = 0.5)
shade(mu_pi, sp_seq, xpd = TRUE)

# predictions for site = 2
mu <- link(m2.4, data = data.frame(species = sp_seq, leaf = 2))
mu_mean <- apply(mu, 2, mean)
mu_pi <- apply(mu, 2, PI, prob = 0.89)
lines(sp_seq, mu_mean, lty = 3, lwd = 0.5)
shade(mu_pi, sp_seq, xpd = TRUE)

# predictions for site = 3
mu <- link(m2.4, data = data.frame(species = sp_seq, leaf = 3))
mu_mean <- apply(mu, 2, mean)
mu_pi <- apply(mu, 2, PI, prob = 0.89)
lines(sp_seq, mu_mean, lty = 0, lwd = 0.5)
shade(mu_pi, sp_seq, xpd = TRUE)
```

- there is no difference between `site`, and minute differences between the 2 `species`, according to the model

```{r}
mu_mean <- apply(mu, 2, mean)
mu_HPDI <- apply(mu, 2, HPDI, prob = 0.89)

sim.herb <- sim(m2.4, data = list(species = sp_seq, leaf = leaf_seq))
str(sim.herb)

herb.PI <- apply(sim.herb, 2, PI, prob = 0.89)

# plot
plot(herb ~ species, data = d_town_slim, col = col.alpha(rangi2, 0.5))
lines(sp_seq, mu_mean)
shade(mu_HPDI, sp_seq)
shade(herb.PI, sp_seq)
```

```{r}
prior_m2.4 <- extract.prior(m2.4)
```


```{r}
mu_sp1 <- link(m2.4, post = prior, data = data.frame(species = 1, leaf = leaf_seq))
mu_sp2 <- link(m2.4, post = prior, data = data.frame(species = 2, leaf = leaf_seq))

mu_sp1_mu <- apply(mu_sp1, 2, mean)
mu_sp1_ci <- apply(mu_sp1, 2, PI, prob =  0.97)
mu_sp2_mu <- apply(mu_sp2, 2, mean)
mu_sp2_ci <- apply(mu_sp2, 2, PI, prob =  0.97)

plot(NULL, xlim = c(1, 2), ylim = c(-5, 50),
     xlab = "species", ylab = "herbivory damage")
lines(mu_sp1_mu, mu_sp2_mu)
shade(mu_sp1_ci, sp_seq)
shade(mu_sp2_ci, sp_seq)
```


On the natural scale (code incomplete)

```{r}
plot(d_town)
```


Plot and compare posterior means from `m3` and `m4`. 

The function `apply` applies a function to a matrix

```{r}
# extract sample
post_m8 <- extract.samples(m8)

d_town$herb.est <- logistic(apply(post_m8$b, 2, mean))
```

```{r testing out code}
precis(post_m8$b)
```


```{r}
# extract sample
post_m8 <- extract.samples(m8)

# compute median intercept for each tamp
# also transform to prob with log
d_town$avg.est <- apply(post_m8$b, 2, mean)

# display raw proportions of herbivory for each individual
plot(d_town$avg, ylim = c(0, 1), pch = 16, xaxt = "n",
     xlab = "individual", ylab = "herbivory proportion", col = rangi2)
axis(1, at = c(0, 40, 80, 100), labels = c(0, 40, 80, 100))

# overlay posterior means
points(d_town$avg.est)

# mark posterior mean probability across all individuals
abline(h = mean(inv_logit(post_m8$a_bar)), lty = 2)
```



Sample the posterior distribution from `m4`. Still working on this code as well.

```{r}
post <- extract.samples(m8)
d_town$herb.est <- logistic(apply(post$b, 2, mean))

plot(d_town$herb.est, col = rangi2, pch = 1,
     xlab = "Individual", ylab = "Herbivory %")
points(d_town$herb.est, pch = 1)

abline(h = mean(inv_logit(post$a_bar)), lty = 2)
```

## Sampling from posterior?


```{r}
# show first 100 populations of the posterior
plot(NULL, xlim = c(-3, 4), ylim = c(0, 0.35),
     xlab = "log-odds ?", ylab = "?")

for(i in 1:100)
   curve(dnorm(x, post$a_bar[i], post$sigma[i]), add = TRUE,
         col = col.alpha("black", 0.2))

# sample 8000 imaginary individuals from post distribution
sim_ind <- rnorm(8000, post$a_bar, post$sigma)

# transform to probability and visualize
dens(inv_logit(sim_ind), lwd = 2, adj = 0.1)
```



Sample the posterior distribution from `m3`. Also working on this code still.

post <- extract.samples(m3)

```{r}
# show first 100 populations of the posterior
plot(NULL, xlim = c(-3, 4), ylim = c(0, 0.35),
     xlab = "log-odds ?", ylab = "?")

for(i in 1:100)
   curve(dnorm(x, post$a_bar[i], post$sigma[i]), add = TRUE,
         col = col.alpha("black", 0.2))

# sample 8000 imaginary individuals from post distribution
sim_ind <- rnorm(8000, post$a_bar, post$sigma)

# transform to probability and visualize
dens(inv_logit(sim_ind), lwd = 2, adj = 0.1)
```


I don't think there is shrinkage here?



# Questionable code

Here is my data:
```{r}
library(readr)
d_town <- read_csv("d_town.csv")
mean(d_town$herb)
```

The `herb` column is a measure of herbivore percent. The average herbivore damage is 0.92%, or close to 1%.

# Priors

I test some priors:

```{r}
set.seed(1)
N <- 100

a <- rlnorm(N, 1, 0.5) # global average
b <- rlnorm(N, 0, 1.5) # species
g <- rlnorm(N, 2, 5) # site

plot(x = "species", y = "herb", data = d_town,
     xlim = range(d_town$species),
     ylim = c(-5, 100),
     xlab = "species", ylab = "herb")
for(i in 1:N)
   curve(a[i] + b[i]*x + g[i],
         from = 1, to = 2, add = T,
         col = col.alpha("black", 0.2))

```




Slim list of variables. 
```{r}
d_town_slim <- list(
   herb = d_town$herb,
   species = as.integer(d_town$species),
   site = d_town$site,
   leaf = d_town$leaf
)

str(d_town_slim)
```




Here's a model:

```{r}
m2.2 <- ulam(
   alist(
      herb ~ dbinom(100, p),
      logit(p) <- a + b[species] + d[leaf],
      a ~ dnorm(1, 0.5),
      b[species] ~ dnorm(0, 1.5),
      d[leaf] ~ dnorm(2, 5)
   ), data = d_town_slim, chains = 4, cores = 4,
   log_lik = TRUE
)
```


I used these priors because they are weakly informative, but do the priors that I test earlier translate to the priors I used here? I am confused by the `logit` and `rlnorm` part - do they do the same thing?




# Mixture model


The way I am thinking about it is that there are 2 ways to get 0 herbivory (resulting in the zero inflation of the dataset): either an individual did not have any herbivore damage because there were no herbivores; or, a leaf did not get herbivorized but the other 2 leaves on the individual were damaged.

First, I created a column `binom_herb` where 0 = no herbivory on any leaves and 1 = herbivory on at least one leaf. 

```{r}
d_town %>% 
   spread(leaf, herb, sep = "") %>% 
   mutate(binom_herb = ifelse((leaf1 + leaf2 + leaf3 == 0), 0, 1)) %>% 
   pivot_longer(
      cols = leaf1:leaf3,
      names_to = "leaf",
      values_to = "herb"
   )
```


So, here is the code for the model

```{r}
m8.2 <- ulam(
   alist(
      herb ~ dzipois(p_zero, 100, lambda),
      logit(lambda) <- a,
      logit(p_zero) <- binom_herb[species],
      
      # priors
      a ~ dnorm(1, 0.5),
      binom_herb[species] ~ dnorm(0, 1.5)
   ), data = d_town_slim, log_lik = TRUE
)

precis(m8.2, depth = 2)
```

So, the MAP estiamtes should be:

```{r}
inv_logit(1.46) 
inv_logit(0.74)

exp(0.99)
```

So, there is a 81% chance that species 1 will not herbivorized, and a lower value of 68% that species 2 will not herbivorized. If there are herbivores present (as indicated by herbivore damage on at least one of the three leaves), an individual faces about 2.69% herbivore damage.


I really am not confident with this code, the priors I chose, or my interpretation of it at all. Also, if I were to add parameters (e.g., `site`) into the regression, which regression do I add it into?

```{r}
m8 <- ulam(
   alist(
      herb ~ dzipois(p_zero, lambda),
      logit(lambda) <- a + b[species],
      logit(p_zero) <- herb_zero,
      
      # priors
      a ~ dnorm(1, 0.5),
      b[species] ~ dnorm(0, 1.5),
      herb_zero ~ dnorm(2, 5)
   ), data = d_town_slim, log_lik = TRUE
)


m8.1 <- ulam(
   alist(
      herb ~ dzipois(p_zero, 100, lambda),
      logit(lambda) <- a + b[species],
      logit(p_zero) <- binom_herb[species],
      
      # priors
      a ~ dnorm(1, 0.5),
      b[species] ~ dnorm(0, 1.5),
      binom_herb[species] ~ dnorm(2, 5)
   ), data = d_town_slim, log_lik = TRUE, chains = 4, cores = 4,
)

m8.2 <- ulam(
   alist(
      herb ~ dzipois(p_zero, 100, lambda),
      logit(lambda) <- a + b[species] + g[site] + d[leaf],
      logit(p_zero) <- binom_herb[species],
      
      # priors
      binom_herb[species] ~ dnorm(2, 5),
      a ~ dnorm(1, 0.5),
      b[species] ~ dnorm(0, 1.5),
      g[site] ~ dnorm(2, 5),
      d[leaf] ~ dnorm(2, 5)
   ), data = d_town_slim, log_lik = TRUE, chains = 4, cores = 4,
)

precis(m8, depth = 2)
precis(m8.2, depth = 2)
```


```{r}
inv_logit(1.72)
```
























# NOTES

Note on multilevel models

- think abut the causal model first
- begin with "empty" model with varying intercepts on relevant clusters
- standardize predictors
- use regularizing priors (simulate)
- add in predictors and vary their slopes
- can drop varying effects wwith tiny sigmas
- consider two sorts of posterior prediction
   - same units: what happened in these data?
   - new units: what might we expect for new units?
- your knowledge of domain trumps all





Variance-covariance matrix
```{r}
vcov(m1)
```

Vector of variances for parameters
```{r}
diag(vcov(m1))

```

Correlation matrix
```{r}
cov2cor(vcov(m1))
```







- site as fixed effect (eventually multilevel model)
- ignore town (not enough data)
      * or only for one species
- gLM with multivariable models
- write out models with priors
      - e.g., y ~ N(mu, sigma)
- fit in quap or Stan
- some description of how to diagnose model
      - e.g., posterior predictions, etc.































